---
title: "HWK 1 - Zambrano"
output: word_document
---
#Kyle Hafey
#4.21.2016
##Question 1
```{r}
#Do the Arsenic/fish problem with new parameters
#Ho: Theta <=.09
#ha: Theta>.09
#log normal random variabel with parameters:
#Mean of log(theta)
logmean = -2.12
#Sd of log(theta)
logsd = 1.34
#sample
stdev = (.028)/(30^.5)
xbar = .052
n = 20000

#Do simulation and calculate probabilities
theta = rlnorm(n,logmean,logsd)
xsim = c()
for(i in 1:n){
  xsim[i] = rlnorm(1, theta[i], stdev)
}
simulation1 = data.frame(theta,xsim)
#View(simulation1)

safelevel = .09
#Calcualte simulations below the safe level
ho = ifelse(theta <= safelevel,1,0)
ho = sum(ho)
#calculate probability of being below the safe level
prho = ho/n
#Calcualte the simulations above the safe level
ha = ifelse(theta > safelevel,1,0)
ha = sum(ha)
#calculate probabilirty of being above the safe level
prha = ha/n

#Calculate probabilities from the realized value of xbar.
xbar = .052 #From earlier
#Use epsilon of .10 from your example
epsilon = .10
#Calculate interval
lowerbound = xbar*(1-epsilon)
upperbound = xbar*(1+epsilon)
meanininterval = numeric(n)
table = cbind(simulation1,meanininterval)
#View(table)
for(i in 1:n){
  if(table$theta[i] >= lowerbound & table$theta[i]<= upperbound){
    table$meanininterval[i] = theta[i]
  }
}
marker = ifelse(table$meanininterval > 0,1,0)
thetagivendata = sum(marker)
sho = ifelse(table$theta < safelevel,1,0)
trueho = sum(sho)
trueha = n-trueho
hodata = ifelse(table$meanininterval < safelevel & table$meanininterval > 0,1,0)
hoanddata = sum(hodata)
haanddata = thetagivendata-hoanddata
pdatagivenho = hoanddata/trueho
pdatagivenha = haanddata/trueha

#Calculate Prior Odds, Bayes Factor, and Posterior Odds.
priorodds = prha/prho
bayesfactor = pdatagivenha/pdatagivenho
posteriorodds = priorodds*bayesfactor
ph1givendata = posteriorodds/(1+posteriorodds)
#Present results
output = data.frame("Prior_Odds" = priorodds, "Bayes_Factor" = bayesfactor, "Posterior_Odds" = posteriorodds)
(output)
```
From this example, we can see that the prior odds are 1.39 to 1 that the mercury levels in the water are above the safe level of 0.09. We simulated these results by using a log normal distribution and estimation methods. The bayes factor and the posterior odds are equal to 0. In this case, we can change the bayes factor and the posterior odds by changing the value of epsilon that sets the interval. The 0 bayes factor and 0 posterior odds tells us that the probability of our simulated date being produced under the alternative hypothesis is zero. We should  consider changing the value of epsilon that creates our interval. This will allow us to improve the odds of actually observing this simulation. 


##Question 2
###Part a
```{r}
#Ho: theta = 0
#Ha:= theta~U(0,1)
#Create base parameters
rand = runif(n)
sdd = 5/sqrt(100)
#Create simulation
theta = ifelse(rand<.5,0,runif(n))
xsim = rnorm(n,theta,sdd)
#Compute test statistic/Zscore
teststat = (xsim-theta)/sdd
simulation2 = data.frame(theta,xsim,teststat)

#part a
#Subset data with those that meet pval = .05 criteria
pval = simulation2[which(teststat >=1.63 & teststat <1.67),]
#View(pval)
nnn = length(pval[,1])
noeffect = ifelse(pval[,1] ==0,1,0)
noeffectt = sum(noeffect)
effectt = nnn-noeffectt
prnoeffect = noeffectt/nnn
preffect = effectt/nnn
results = data.frame(prnoeffect,preffect)
View(results)
```
From this simulation, 44% will have no effect and 56% will have an effect. These results are subject to change based on the simulation, but the results should be presented int he table. This means that 46% of the results that has a p value of .05 actually had a negligible effect(no effect). this simulation is questionable.  

###Part b
```{r}
#calculating error types?
#Subset data into null and alt by subseting the theta values
null = simulation2[which(simulation2[,1]==0),]
nnull = length(null[,1])
alt = simulation2[which(simulation2[,1]!=0),]
nalt = length(alt[,1])  
#Only allow values that we reject for both the null and alternative hypothesis
rejectnull = null[which(null[,3] > 1.65),]
nrejectnull = length(rejectnull[,1])
rejectalt = alt[which(alt[,3] > 1.65),]
nrejectalt = length(rejectalt[,1])
#Now, Calculate the pre experimental rejection ratio
#Power should be about .05 because the significance value is .05
power = nrejectalt/nalt
#rpre = 1-power/alpha
rpre = (1-power)/.05
(rpre)
```
The power is about .05 which makes sense because we are analyzing it with a p-value of .05. The pre experimental rejection ratio is around 19, but this can change due to the simulation aspect of the estimation. It should be around 18 or 19. 


##Question 3
```{r}
pval3 = simulation2[which(simulation2[,3] >=2.326 & simulation2[,3] <2.366),]
xsim2 = rnorm(n, simulation2[,1],sdd)
teststat2 = (xsim2-simulation2[,1])/sdd
simulation3 = data.frame(theta,xsim2,teststat2)
#View(simulation3)
#Subset again
probability = simulation2[which(simulation2[,3] >=2.346 & simulation2[,3] >2.346),]
nprobability = length(probability[,1])
Results = nprobability/n
(Results)
```
I estimate the probability that 1 sided p-value from the second experiment will be 1% or smaller, given that the first measurement had a p-value of p=.01 to be around 1%. Depending on the simulation, the value can change, but it will be very small. This means the researchers results will be extremely hard to replicate. It will only be replicated successfully about 1% of the time. 

















